From OpenACC to OpenMP 4: \\ Toward Automatic Translation

### 2016-06-13 paper.tex
Add 4.2 Sequential Private/Reduction Removal

\documentclass{sig-alternate-05-2015}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{xspace}
\usepackage[table]{xcolor}
\hyphenpenalty=750
\newcommand\todo[1]{\textcolor{red}{** #1}}
\newcommand{\pooda}{\textsc{POOD}\xspace}

@@

The translation algorithm is as follows.  The input is a C source file
containing OpenACC directives.  The output is the same file with OpenMP
directives substituted.
\begin{enumerate}
\item If any of the following occur, warn the user; these structures will
not be automatically converted:
\begin{itemize}
\item The header file \texttt{openacc.h} is included.
\item The macro \texttt{\_OPENACC} is used.
\item Functions in the OpenACC runtime library are called or referenced.
\item The \texttt{kernels} directive is used.
\end{itemize}
\item Prepare parallel loop nests using the algorithm described in
\S\ref{sec:loops} below.
\item Apply the translation rules described below in \S\ref{sec:directives} to
convert OpenACC directives to OpenMP directives.
\end{enumerate}

@@

@@

\newpage

@@

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
\end{document}

>>>>

\documentclass{sig-alternate-05-2015}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{xspace}
\usepackage[table]{xcolor}
\usepackage{balance} % for \balance
\hyphenpenalty=750
\newcommand\todo[1]{\textcolor{red}{** #1}}
\newcommand{\pooda}{\textsc{POOD}\xspace}

@@

The translation algorithm is as follows.  The input is a C source file
containing OpenACC directives.  The output is the same file with OpenMP
directives substituted.
\begin{enumerate}
\item If any of the following occur, warn the user; these structures will
not be automatically converted:
\begin{itemize}
\item The header file \texttt{openacc.h} is included.
\item The macro \texttt{\_OPENACC} is used.
\item Functions in the OpenACC runtime library are called or referenced.
\item The \texttt{kernels} directive is used.
\end{itemize}
\item Prepare parallel loop nests using the algorithm described in
\S\ref{sec:loops} below.
\item Remove \emph{private} and \emph{reduction} clauses from sequential
loops as described in \S\ref{sec:privred} below.
\item Apply the translation rules described below in \S\ref{sec:directives} to
convert OpenACC directives to OpenMP directives.
\end{enumerate}

@@

\subsection{Sequential Private/Reduction Removal}
\label{sec:privred}

\begin{figure*}[t]
\begin{center}
{\footnotesize
% (a) Original
\begin{minipage}{.5\textwidth}
\texttt{%
int x, sum = 100; \\
\textcolor{blue}{\bfseries \#pragma acc loop seq private(x) reduction(+:sum)} \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} x = 100; \\
\hspace*{2em} sum += x; \\
\}
}
\end{minipage}
\hspace*{.07\textwidth}
% (b) Original
\begin{minipage}{.36\textwidth}
\texttt{%
int x; \\
\textcolor{blue}{\bfseries \#pragma acc loop seq reduction(+:x)} \\
for (int i = 0; i < 5; i++) \\
\hspace*{2em} x = 1;
}
\end{minipage} \\
% Arrows
\begin{minipage}{.5\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
\end{minipage}
\hspace*{.07\textwidth}
\begin{minipage}{.36\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
\end{minipage} \\
% (a) Transformed
\begin{minipage}{.5\textwidth}
\texttt{%
int x, sum = 100; \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} \textcolor{blue}{\bfseries int x;} \\
\hspace*{2em} x = 100; \\
\hspace*{2em} sum += x; \\
\}
}
\end{minipage}
\hspace*{.07\textwidth}
% (b) Transformed
\begin{minipage}{.36\textwidth}
\texttt{%
int x; \\
\textcolor{blue}{\bfseries int sum = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (1)} \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} \textcolor{blue}{\bfseries int x = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (2)} \\
\hspace*{2em} x = 1; \\
\hspace*{2em} \textcolor{blue}{\bfseries sum += x;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (3)} \\
\} \\
\textcolor{blue}{\bfseries x = sum;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (4)}
}
\end{minipage} \\
% Labels
\begin{minipage}{.5\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} (a)}
\end{minipage}
\hspace*{.07\textwidth}
\begin{minipage}{.36\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} (b)}
\end{minipage} \\
}
\end{center}
\caption{Elimination of \emph{private} and \emph{reduction} clauses on a sequential loop.}
\label{fig:privred}
\end{figure*}

After a loop nest has been prepared, there will be a gang loop, a vector loop
(possibly the same loop), and zero or more sequential loops.  OpenACC has a
directive for sequential loops (\texttt{acc loop seq}), which permits
\emph{private} and \emph{reduction} clauses.  However, OpenMP does not have a
sequential loop directive.  This means that these directives---and more
importantly, \emph{private} and \emph{reduction} clauses---must be removed from
sequential loops in a way that will preserve the semantics of the
\emph{private} and \emph{reduction} clauses.

To remove a \emph{private} clause, we can simply re-declare the privatized
variable in the loop body, as shown in Figure~\ref{fig:privred}(a) (noting that
this features requires C99 support, which is available in all modern C
compilers).  This declaration shadows the declaration in the outer scope,
which gives the intended semantics of the \emph{private} clause.

To illustrate removing a \emph{reduction} clause, consider as a concrete example
the clause \texttt{reduction(+:x)}.  There are two cases.

\textit{Case~1}.  This is perhaps the most commonly occurring use of the
\emph{reduction} clause: If the only accesses to \texttt{x} in the loop
body are in statements of the form \texttt{x++} or \texttt{x\ +=\
}\emph{expression} (where \texttt{x} is not assigned in \emph{expression}),
then the reduction clause can be removed with no change in semantics.  This is
the case in Figure~\ref{fig:privred}(a), for example.

\textit{Case~2}.  When the above condition does not hold, it is still possible
to remove a \emph{reduction} clause.  However, preserving the semantics of the
\emph{reduction} clause requires a more complex source code transformation,
illustrated in Figure~\ref{fig:privred}(b).
(1)~A ``fresh'' variable (\texttt{sum}) is declared to accumulate the result.
(2)~Since \texttt{x} is the reduction variable, a local variable \texttt{x}
is declared in the loop body with the same type as the shadowed variable
\texttt{x}, and it is initialized to the initialization value for a sum
reduction (\texttt{0}).
(3)~At the end of the loop body, the local \texttt{x} is added to the result
(\texttt{sum}).
(4)~After the loop, the accumulated result is stored in the original
\texttt{x}.

@@

@@

\balance
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
\end{document}
